---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
str(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
library(tidyr)
library(readr)
library(dplyr)
deaths <- av %>% pivot_longer(cols = c(`Death1`,`Death2`,`Death3`,`Death4`,`Death5`),
                                names_to = "Time",
                                names_transform = parse_number,
                                values_to = "Death",
                                values_transform = tolower) 
  
returns <- av %>% pivot_longer(cols = c(`Return1`,`Return2`,`Return3`,`Return4`,`Return5`),
                                names_to = "Time",
                                names_transform = parse_number,
                                values_to = "Return",
                                values_transform = tolower) 


```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

## Ishan Nichols:

Quote: "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team."


### Include the code
## Bhavika Rathi
> Quote the statement you are planning to fact-check.
There's a 2-in-3 chance that a member of the Avengers returned from their first death.
# Look at Time = 1 deaths only
first_deaths <- deaths %>%
  filter(toupper(Death) == "YES", Time == 1)

first_returns <- returns %>%
  filter(toupper(Return) == "YES", Time == 1)

total_first_deaths <- nrow(first_deaths)
total_first_deaths

returned_after_first <- first_deaths %>%
  inner_join(first_returns, by = "Name.Alias") %>%
  nrow()
returned_after_first

return_rate <- returned_after_first / total_first_deaths
return_rate

cat(sprintf(
  "Died first time: %d\nReturned after first death: %d\nReturn rate: %.1f%%\nArticle claims: ~66.7%%\n",
  total_first_deaths, returned_after_first, return_rate * 100
))
The analysis shows that all avengers who die for the first time, 2 out of 3 did return. This claim is supported by the data. 

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.




First, we will check to make sure the dataset has a matching number of 173 unique Avengers.

```{r}
str(av)
sum(duplicated(av$URL) == TRUE)

```

Since there are 173 observations and no duplicates, we can confirm that the they did analyze 173 Avengers. Now we will check the number of Avengers who have at least one death.
```{r}
sum(av$Death1 == "YES")
69/173
```
There are 69 Avengers who have at least one death, and 69 of of 173 is 39.88 which rounds to 40%. Therefore, we can confirm that the claim made by the quote is accurate.




